### 2018-05-18 paper.Rmd
Add files via upload

---
title: 'The effects of symbols and information in government communication on citizens  trust in policy: Evidence from a survey experiment'
author: '**Saar Alon-Barkat**, PhD candidate; the Hebrew University of Jerusalem;   saar.barkat@mail.huji.ac.il'
date: " "
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
urlcolor: blue
---

Preliminary draft. Last edited at `r Sys.Date()`

```{r , include=FALSE, echo=FALSE}
#load("SVIVA_R_ENV.RData")
source("SVIVA_exp2_dm_03.R")

SVIVA2_01_haifa = filter(SVIVA2_01,AREA==1)
SVIVA2_01_center = filter(SVIVA2_01,AREA==0)
```

```{r silent-packages, echo = FALSE, eval = TRUE, message=FALSE, include = FALSE}
library(knitr)
library(tidyverse)
library(stargazer)
library(car)
library(broom)
library(kableExtra)
```

```{r,echo=FALSE,warning=FALSE,message=FALSE}
N_raw = nrow(SVIVA2_raw_00)
N_0=SVIVA2_00%>%nrow()
N_1=SVIVA2_01%>%nrow()

filter_IP = N_0-(SVIVA2_00 %>%
  distinct(IP,.keep_all=TRUE) %>% nrow())
filter_age = N_0-(SVIVA2_00 %>%
  filter(!(AGE%in% 1:17))%>% nrow())
filter_IMC = N_0-(SVIVA2_00 %>%
  filter(IMC==1)%>% nrow())
filter_time = N_0-(SVIVA2_00 %>%
  filter(TIMER_total>=3,TIMER_total<=30) %>% nrow())

N_1_haifa = filter(SVIVA2_01,AREA==1) %>% nrow()
N_1_center = filter(SVIVA2_01,AREA==0) %>% nrow()

```

<br>

---

# Abstract

*Effective public communications have been proposed as a remedy for citizens' distrust in government. The studies that explored the effectiveness of government communications have focused mainly on the extent to which they provide substantive information about bureaucracies' performance and policies. Recently, a few studies pointed to the effect of symbolic elements, entangled in these communications (e.g. logos, images and celebrities). Still, we lack theoretical understanding of the different mechanisms underlying citizens' responses to information and symbolic elements, and accordingly - the interaction between them. Building on the Elaboration Likelihood Model (ELM) from social psychology, I hypothesize that symbolic elements have a greater effect on citizens' trust the less citizens perceive the issue as personally relevant. Also, I hypothesize that symbolic elements are likely to lead citizens to dedicate less attention to the substantive information in the communication, and accordingly, to increase their trust in logically-unpersuasive policy plans. I put these hypotheses to test through a survey experiment, where I examine Israeli citizens' trust in the Environment Protection Ministry's policies for reducing the air pollution in a specific area (the Haifa-Bay) and reducing domestic waste. The policies are presented to subjects under different conditions of substantive information (logically persuasive vs. unpersuasive policies) and symbolic elements (logo, color, and celebrities endorsers). I operationalize personal relevance by comparing between citizens from the polluted Haifa-Bay area and others. The results do not support the hypothesis that personal relevance attenuates the effect of symbols. Rather, they suggest that symbolic elements increase citizens' trust in ill-perceived policies, by distracting them from scrutinizing their content.*

---

<br>

# Introduction

Public administration scholars are highly concerned about the implications of citizens’ low trust in government, and their undervaluation of public-service performance. One of the main remedies which has been proposed for this problem is greater investment in direct government communications with citizens. Congruently, over the past generation, economically-developed democratic countries have witnessed an increase in government public communications, specifically through advertising, social media, and e-government platforms [@deacon_2001; @eshuis_2012; @marland_2017; @mergel_2013; @mickoleit_2014]. In the US, for example, federal government agencies spend approximately 1.5 Billion dollars annually on contracts related to public relations and advertising activities in traditional and digital media [@unitedstatesgovernmentaccountabilityoffice_2017].

The public administration research on the effect of government public communications on citizens attitudes has focused mostly on the *informative* aspects of these communications, whereby government organizations provide citizens with evidence about their performance and details and explanations about their policies and actions [e.g. @marvel_2015; @im_2014; @porumbescu_2017; @porumbescu_2016; @porumbescu_2017a; @porumbescu_2017b; @grimmelikhuijsen_2015; @ho_2017]. Public communications are expected to increase citizens' trust insofar as they provide citizens with favorable information on organizations' performance and logically persuasive explanations for their policies. At the same time, experiment-based studies that explored the responses of citizens to substantive information revealed cognitive biases that distort their processing of information in communications, and undermine its effectiveness [@baekgaard_2016; @grimmelikhuijsen_2013; @grimmelikhuijsen_2014; @hvidman_2016; @james_2011; @james_2014; @james_2017; @marvel_2015; @olsen_2017; @james_2017a; @grimmelikhuijsen_2017].

Recently, a few studies pointed to the additional effect of *symbolic* elements that are entangled in government communications, such as  agencies' names and brand logos, figures, images and celebrity endorsements [@alonbarkat_2017; @karens_2016; @teodoro_2018]. These studies demonstrated that government organizations may cause citizens to view them more favorably through familiar and well-designed symbolic elements that evoke positive associations and emotions. Still, we lack a theory and robust empirical evidence of the exact mechanisms through which these symbols increase citizens trust. Additionally, these previous studies did not examine the interaction between the symbolic elements and the substantive content of information in public communications.

This research gap has normative significance. Whereas the responsiveness of citizens to substantive information is perceived favorably from a normative view of democratic theory, the potential effect of symbolic elements can be problematic. One can argue that shaping citizens' perceptions of governments by evoking associations and emotions stands in tension with the normative expectation that citizens' judgments should be based on their critical thinking. Mainly, we should be concerned about the possibility that symbolic communication may elicit undue trust in poorly performing organizations and in ill-conceived policies, and enable public organizations to escape justified public criticism [@alonbarkat_2017]. Therefore, this study focuses on the following questions: *What are the psychological mechanisms by which symbolic elements in governemnt communications affect citizens' trust in government organizations and their policies?* Accordingly, *under what conditions do symbolic elements  increase citizens' trust in governemnt policy?* And specifically, *does the effect of symbolic elements depend on the substantive content of communications?* *Do symbols offset the rational, negative effect of unfavorable or logically unpersuasive information?*

To address these research questions, I draw upon the social psychology Elaboration Likelihood Model (ELM). ELM provides a comprehensive theoretical framework for the varying psychological processes underlying individuals' responses to persuasive communication, and their consequences for attitude change  [@petty_1986; @petty_2011]. Based on ELM, I suggest that symbolic elements in public communications can affect citizens' trust via two mechanisms: *First*, they may evoke positive feelings and associations that are transferred from the symbols to the relevant government entities and policies, and thereby increase citizens' trust in them. I expect that this effect is attenuated the more citizens perceive the communications as personally relevant, because they are more motivated to think about their messages. *Second*, symbolic elements may increase citizens' trust in government entities and policies by causing them to disregard negative of logically unpersuasive content that would have otherwise decreased their trust.

To empirically test these theoretical expectations, I conducted a large randomized survey experiment among circa `r SVIVA2_01%>%nrow() %>% round(-1)` Israeli citizens. The experiment focuses on the empirical case of the Israeli Environment Protection Ministry. I explore citizens' responses to public communications about two policy plans of the Ministry, and specifically the extent to which their trust in these policies are affected by the substantive information of the communications (logically persuasive vs. unpersuasive policy plans) and by symbolic elements in them (real symbols vs. fake symbols and no symbols conditions). In addition, I examine the moderating role of the perceived personal relevance of these policies.

I find that the symbolic elements increased citizens' trust in these policies. In line with my theoretical expectations, the effect was greater among those who saw the logically unpersuasive policy plans, who were distracted by the symbols and thus were less inclined to think about its content. I did not find support for the hypothesis that the effect of the symbolic elements is attenuated by the perceived personal relevance of the policy issue. I discuss the theoretical and normative implications of these findings.

<br>

# Symbolic elements in government communications and their shaping of citizens' attitudes: what do we know?

Symbols are fundamental for almost any form of human interaction, as acknowledged by early scholars of social science [@blumer_1986]. A symbol can be broadly defined as *"any object used by human beings to index meanings that are not inherent in, nor discernible from, the object itself"* [@elder_1983, pp. 28-29]. My focus is on those symbols that are strategically used by government organizations in their public communications, and are designed to distinguish organizations and to evoke associations and emotions that stimulate audiences’ positive affect towards organizations and their operations. Prominent examples are unique logos, colors associated with specific meanings, icons, images, and the use of celebrities as endorsers. These elements are widely used by businesses and political candidates to attract targeted consumers and voters. Yet, they can be also found in the communications of government bureaucracies with citizens, including their mass-media advertising campaigns, social media accounts, internet pages, official publications, signboards in public places and personal mail notifications to citizens.

By comparison with the abundant research on citizens’ responses to informative elements in public communications [for reviews see: @cucciniello_2016; @grimmelikhuijsen_2017; @james_2017], public administration literature has devoted relatively little attention to the effect of symbolic elements. Recently, however, a few experimental studies tackled the issue. @marvel_2015 analyzed US citizens' responses to a television advertisement of the public Postal Service, which involved various symbolic elements, including the organization's logo, workers' images and background music. His study shows that participants who watched the advertisement evaluated the organization's performance more positively. @teodoro_2018 suggested that the name of US federal agencies not only represents information about their activities, but may also carry specific symbolic meaning and affect that can shape citizens' support for government policy (depending on citizens' political party identification). These two studies provide examples for the role of symbols in shaping citizens' attitudes. Still, they did not effectively disentangle the effects of the symbolic and informational aspects of the studied public communications.

A study by @karens_2016 did examine the distinct effect of symbols in communications on citizens' trust. The scholars surveyed university students from three European countries and examined their trust in two policy plans of the European Commission, and it's shaping by the appearance of the organization's logo in the communication of the policy plans. They found that including the Commission’s logo significantly increased citizens' trust in these policies. Their results show an impressive effect size equivalent to 0.7 standard deviations. [^footnote_karens]

[^footnote_karens]: The effect size they found was smaller with regard to the "air quality" policy compared with the "digital skills" policy which were presented (Cohen's d = 0.7 vs. 0.6). Also, the effect was smaller among subjects from the Netherlands, compared with those from Belgium and Poland (Cohen's d = 0.5 vs. 0.7).

@alonbarkat_2017 examined Israeli citizens' trust in the Israel Electricity Corporation, a state owned company, and it's shaping by their exposure to comic cartoon figures that were routinely presented in the organization's public campaigns. They compared between citizens who were personally exposed to prolonged power outages, and others who were not. They found that the symbolic element had a positive effect on citizens' trust in the organization, including among those who personally experienced poor performance (a modest effect size equivalent to 0.25 standard deviations). Building on social psychology theory and marketing research, they suggested that the symbols enhance trust by inducing an unconscious transfer of affect from the symbols to the organizations and policies associated with them (evaluative conditioning). They proposed that the effectiveness of government symbolic elements depends on their familiarity, and aesthetic design, as well as on the ambiguity of the assessed government quality.

The aforementioned studies tend to suggest that public communications increase citizens' trust in government organizations and their policies, not only by conveying valuable substantive information, but also by incorporating symbolic elements that evoke positive associations and emotions. Moreover, the findings of @alonbarkat_2017 suggest that incorporating symbolic elements in public communications can unduly compensate for poor government performance, thus pointing to their potential detrimental implications.

Even though these studies substantiate a link between the appearance of symbolic elements in communications and citizens' favorable attitudes, they did not directly investigate the causal mechanisms that underlie this effect. Therefore, we still lack understanding of the circumstances under which symbols are likely to be more/less effective in shaping citizens' trust. Additionally, these studies neither theorized nor empirically tested the interaction between symbolic elements and citizens’ processing of substantive information in communications. We do not know whether well-designed symbols reinforce, or offset, citizens' rational responses to information. Specifically, as explained above, it is essential to assess the possibility that symbols can unduly compensate for unfavorable or logically unpersuasive information that should yield negative responses.

To address this theoretical gap in public administration research, I propose a nuanced theoretical framework for the effects of symbols in communications and their interaction with information, building on social psychology theory on persuasive communication.

<br>

# The Elaboration Likelihood Model and the various mechanisms of persuasion by elements in communications

The social psychology Elaboration Likelihood Model (ELM) provides a comprehensive theoretical framework for investigating peoples' varying responses to "persuasive" communications [@petty_1986; @petty_2011]. In addition to its general theoretical application in psychological research, this model has been applied to studies on persuasion in various areas, including commercial advertising [@petty_1983; @swasy_1985; @dotson_2000; @drge_1989; @wheeler_2005; @moore_1987], public health campaigns [@wilson_2007; @angst_2009; @ruiter_2001; @briol_2006], media coverage [@petty_2002; @vanderwurff_2016] and media entertainment [@young_2008; @slater_2002]. Yet, thus far, the relevance of ELM has not been directly examined in relation to citizens' responses to government communications.

According to ELM, peoples' responses to different elements in persuasive communications depends on the amount of thinking that they dedicate to the message content. Persuasion can occur when thinking is "high" or "low", but the processes and consequences are likely to differ in each situation. This range between high and low degrees of thought is referred to as the "elaboration continuum". Along this continuum, there are numerous specific processes of persuasion (or attitude change) that can occur. At the high end of the elaboration continuum, also called the "central route", people are likely to respond by cognitively scrutinizing the arguments central to the merits of the issue. At the low end, also called "the peripheral route" people respond to simple cues automatically and unconsciously through primitive psychological processes that require very little amount of cognitive effort, such as a transfer of affect[^footnote_dual_process_theories]. These different persuasion mechanisms are likely to cause divergent types of attitudinal change. Persuasion through the central route is expected to be more enduring, resistant and predictive of behavior, compared with peripheral route persuasion. Of course, much of the time, people are located in between the two ends of the continuum, and thus persuasion is determined by a mixture of these processes.

[^footnote_dual_process_theories]: The conceptual distinction between the "central and peripheral routes for persuasion" largely resembles the distinctions between heuristic and systematic processing [@chaiken_1980], and systems 1 and 2 [@kahneman_2011].

People at different locations along the elaboration continuum may process the same element in a communication in different ways. Consider for instance a text that contains a series of logical arguments supporting a certain position. Those at the high end of the continuum would regard this text as arguments, and systematically scrutinize it based on its logical persuasive quality (central route processing). Conversely, those at the low end of the continuum are likely to disregard the content of the text, and instead, rely on the number of lines in the text, for example, as a cue for its strength (peripheral route processing). Some elements provide potential cues that lack any logical quality at all, such as the physical attractiveness of the source. These latter elements may affect the attitudes of those at the low end, but are less likely to persuade those at the high end [@petty_1986].

ELM further postulates that the placement of people along the elaboration continuum in any given situation is determined by their motivation and ability to think about the message. The literature proposes a variety of such motivational and ability factors. Most prominently, the more people perceive the issue as personally relevant, the more motivated they are to invest in thinking about it, and accordingly, they more likely they are to use the central route [@petty_1979; @petty_1981; @petty_1983]. Also, studies have demonstrated that the more people are distracted, the more difficult it is  for them to scrutinize the message arguments, a mechanism that has been referred to as "elaboration disruption" [@petty_1976; @petty_1986]. People's location along the elaboration continuum is also determined by their affective state or mood. Positive mood (compared with neutral or negative moods) reduces people's cognitive ability and attenuates their motivation to think, due to their desire to maintain their positive mood and avoid message processing that would depress it [@worth_1987; @mackie_1989; @bless_1990; @schwarz_1991; @petty_2015; but see: @wegener_1995].[^footnote_elm_factors]

[^footnote_elm_factors]: For a thorough review of motivational and ability factors affecting likelihood to elaborate see @petty_1998.

People's motivation and/or ability to think is shaped not only by factors external to the communication, but also by elements in the communication itself. For example, using second-person pronouns in the message (i.e. "you", "your") increases people's perceived personal relevance, and motivation to elaborate [@burnkrant_1989]; Background music in commercials can distract people and prevent them from scrutinizing their content [@park_1986]; The use of humor in television shows stimulates positive mood and reduces people's ability and motivation to scrutinize their substantive arguments [@nabi_2007; @young_2008]. Affecting the amount of thinking can cause attitude change in interaction with elements in the communication that can function as arguments or peripheral cues. Accordingly, decreasing the amount of thinking enhances the effect of peripheral cues, whereas increasing the amount of thinking enhances the effect of the logical arguments.

Finally, some elements in communications can function both as variables that affect the amount of thinking and as potential peripheral cues or arguments. Thus, these elements can shape people's attitudes through more than one causal mechanism. Continuing with the above mentioned examples, background music and humor can function at the same time as factors that reduce elaboration and as peripheral cues. @petty_1986 have suggested that we are more likely to find these complex relations of multiple mechanisms in those situations where people are located in the middle of the elaboration continuum. When people are neither extremely motivated or unmotivated to think, they may rely on different elements in the communications to decide if the message is worth or needs considering.

<br>

# ELM and the roles of symbolic elements and information in government public communications

How may we apply the theoretical framework of ELM to the study of the effects of symbolic elements in government communications with citizens and their interaction with substantive information?

In light of the conceptualization of ELM, substantive information elements in government public communications represent issue-related arguments, which can be scrutinized or processed systematically based on their logical quality (via the central route), at least by those who are sufficiently motivated to do so. Conversely, symbolic elements represent potential peripheral cues, which can affect attitudes through the peripheral route, by evoking positive emotions and associations that are attributed to government organizations and their policies [@alonbarkat_2017]. Accordingly, the effect of symbolic elements on citizens' trust is likely to depend on citizens' motivation to cognitively invest and think about the policy issues. Specifically, I expect that the more citizens perceive the policy issue as personally relevant, the more they are likely to focus on scrutinizing the content of the message, and the less they are likely to rely on the symbolic elements entangled in the communication.

Accordingly, I expect that:

>*H<sub>1</sub> - Symbolic elements in public organizations' communications are likely to increase citizens' trust in the organizations and in the policies presented in the communications.*

>*H<sub>2</sub> - The positive effect of symbolic elements on citizens' trust is likely to be attenuated the more citizens perceive the policy issues in the communications as personally relevant.*

However, we can also derive from ELM a second mechanism by which symbolic elements may increase citizens' trust. In addition to their direct positive effect on trust (via the peripheral route), symbols can decrease the amount of thought citizens dedicate to the communication, and thus prevent them from scrutinizing negative information (e.g. information on poor performance) and weak arguments (e.g. logically unpersuasive description of their policy plans).

Symbolic elements can decrease citizens' ability and motivation to think about the message in several ways. First, they can cause elaboration disruption, by distracting citizens and absorbing their attention and cognitive efforts [@petty_1976; @petty_1986]. Second, they can decrease citizens' motivation and ability to think about the message content by creating a positive affect and placing citizens in a positive mood. As explained above, the postulation that positive mood decreases elaboration has been explained, in part, by a mood maintenance mechanism [@mackie_1989; @bless_1990; @schwarz_1991]. This mechanism is particularly relevant to the context of government communications, in which thinking about the content of the message - government organizations and policies - is likely to yield negative mood among most people [@marvel_2015; @marvel_2015a; @hvidman_2016]. Therefore, when citizens are exposed to symbolic elements that evoke positive feelings, they may seek to avoid thinking about tedious bureaucratic affairs, which would depress their positive mood.

The expectation that symbols will reduce the amount of thinking entails an interaction between symbols and the substantive information in the communications. Symbolic elements are likely to increase citizens' trust insofar as they reduce citizens' inclination to scrutinize information that would have decreased their trust. Accordingly, I expect that:

>*H<sub>3</sub> - The positive effect of symbolic elements on citizens' trust is likely to be enhanced when the communication includes negative or logically unpersuasive information that would have decreased citizens' trust.*

>*H<sub>4</sub> - The abovementioned conditioned effect of symbolic elements on citizens' trust is likely to be mediated by the decrease in the amount of thinking (a moderated mediation hypothesis).*

The theoretical hypotheses are graphically displayed in **Figure 1**.

<br>
**Figure 1: Theoretical model**
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/theoretical_model.png) 
<br>

Analytically, the expectation that symbolic elements decrease message elaboration also entails that symbols may decrease citizens' trust in communications about highly positive performance and persuasive policy plans. On the other hand, one can make the argument that neutralizing the effect of "positive" information with symbols may not actually yield significant differences, for two reasons: First, citizens are more likely to "punish" bureaucracies for publishing negative information than to "praise" them for publishing positive information [@james_2011; @nielsen_2016; @james_2017b]. Congruently, the effect of positive information (versus a medium level) is likely to be relatively small, and hence neutralizing this effect may not make much difference. Second, even when communications include favorable information, symbols can still have a positive effect on trust through the peripheral route (in accordance with my first hypothesis). Given the focus of this study on the interaction between symbols and ill-conceived policies, the research design does not directly addresses this latter interaction, which can be .

<br>

# Methodology

To empirically test the above mentioned hypotheses, I designed a randomized survey experiment, which simulates citizens' responses to government organizations' communications about their policy plans. Such communications can be found in publicized reports, pamphlets, websites, new media pages and ads. Focusing on communication of future policy plans, as opposed to communications on performance of existing policies, enables me to utilize a real-world empirical case without compromising the reliability  of the manipulation of the content. This is because most citizens do not have prior knowledge of organizations' future policy plans, and thus they will be more willing to trust a fictional communication of an organization with regard to its future policy.

<br>

##*The empirical case*

Empirically, I focused on the case of the Israeli Environment Protection Ministry (hereafter: EPM), a national level bureau responsible for the formulation, co-ordination and execution of environmental policy in Israel, both at the national and local levels. I analyzed the effects of informative and symbolic elements in EPM's communications of two policy plans on citizens' trust in these policies, and the extent to which they elaborated on them. In the survey, the communications about the two policy plans, in which I manipulated the informative and symbolic elements, are presented to the participants as sections from the EPM's official publication of its annual work plan. Methodologically, EPM makes a good case study for two main reasons.

*First*, EPM has invested greatly over the past years in public communications. In these various communications, the EPM used an array of symbolic elements, which became widely recognized by the public. Here, I decided to focus on three prominent symbolic elements used in the EPM's public communications, and displayed below in **Figure 2**. The first symbolic element is the unique brand logo of the EPM (image a. in figure 1). This logo consists of pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second symbolic element in which I focused is the green color, which is strongly associated with taking care of the environment. The EPM's logo and the green color can be seen in almost all of the visual communications of the EPM, and they represent key elements of its unique "visual identity" [@vandenbosch_2005]. The third element is celebrity endorsements. Specifically, I selected two high-profile celebrity comedians, *Tal Friedman* and *Ido Rosenblum*, that previously endorsed two salient advertising campaigns of the EPM. The campaign starring the comedian Tal Friedman (image b.) was launched in 2010, and included a series of advertisements in which Friedman was presented with a green grass head, and accordingly the campaign's slogan was: "starting to think green". The campaign starring the comedian Ido Rosenblum (image c.) was launched in 2017, and was focused on reducing the use of disposable bags.[^footnote_campaigns] The two campaigns were highly successful in terms of their recognizability and likability, as suggested by their evaluation reports by the Government advertising agency.

In the course of the survey experiment, I confirmed that the EPA logo and the images of the celebrities from the campaigns are familiar to the participants. `r round((nrow(filter(SVIVA2_01,RECOGNIZE_SVIVA_logo>1))/N_1)*100,0)`% of the participants reported that they recognize EPA logo to some extent and a vast majority of them reported that they recognize the campaign images of Friedman and Rosenblum (`r round((nrow(filter(SVIVA2_01,RECOGNIZE_air_real>1))/N_1)*100,0)`% and `r round((nrow(filter(SVIVA2_01,RECOGNIZE_waste_real>1))/N_1)*100,0)`% , respectively). In robust analyses I further exclude those who do not recognize these symbols.[^footnote_prior_studies]

[^footnote_campaigns]: Multimedia files of these campaigns are available in EPM's website: http://www.sviva.gov.il/InfoServices/ReservoirInfo/ResearchAndPublications/Pages/default.aspx.

[^footnote_prior_studies]: In two prior online survey tests, which I conducted before the experiment, I collected further information about the affect and symbolic associations of these symbols. These results, which are reported in the supplementary appendix [##TBA##], substantiate my empirical assumption that these elements evoke positive emotions Israeli citizens, and that citizens automatically associate them with EPM and its policies. These results also indicate that the images of the celebrity comedians from EPM campaigns are perceived as humoristic, and accordingly, they are expected to position people in a positive mood.

<br>
**Figure 2: symbolic elements of EPM**
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/EPM_symbols.png) 
<font size="0.5">

Notes:

a. EPM logo.
b. Tal Friedman, comedian,  EPM campaign “starting to think green” (2010): "Isn't money growing on trees? Moving to environmental consumption and saving at least 6,000 NIS per year!", "starting to think green", "The Environmental Protection Ministry".
c. Ido Rosenblum, comedian, EPM campaign  “taking every bag seriously” (2017): "The reusable bag is again left at home? A tip from Ido: Hang it on the door handle and you will not leave without it", "Taking every bag seriously", "The Environmental Protection Ministry".

</font>

<br>

The *second* reason for selecting EPM regards its policy plan regarding the pollution in the Haifa-Bay area, which offers a fortunate case for examining the moderating role of perceived personal relevance. The Haifa-Bay area, located on the Mediterranean coast of northern Israel, is characterized by relatively high levels of air pollution, mainly due to the concentration of heavy industry. Beginning in the late 2000s, the Ministry of Environment protection has been implementing policy programs to reduce the high levels of pollution in the area. The uniqueness of this policy issue is rooted in the fact that it represents natural variation in citizens' perceived personal relevance (between those who reside in the polluted area, and others), which is not strongly confounded with differences in personal knowledge, as in most cases [@petty_1983]. The air pollution in the Haifa-bay area  is an intangible, invisible, ambiguous and complicated policy issue. Those who reside in a polluted area, are not likely to have much greater knowledge and understanding of the issue. Accordingly, when they are exposed to a communication about this personally relevant issue, they are more motivated to scrutinize it, yet are not more capable to do so than others. Hence, this case enables to isolate the distinct effect of personal relevance. As explained below, the sample population of the experiment consists of residents of Haifa-bay area, matched with residents of cities at the center of Israel with similar aggregate socio-economic profile.

I supplemented the Haifa-bay air-pollution policy with a second policy plan, which is likely to be perceived as equally relevant for those Haifa-bay residents and others. This comparison further enables me to confirm that differences between the two groups, if found, are explained by differences in the perceived personal relevance of the policy issue. Furthermore, testing the effects of symbolic elements and information on two different policy plans provides a more rigorous test for the theory. Hence, I selected EPM's policy for reducing of domestic waste and increasing recycling, a tangible policy issue which I expect to be perceived as equally personally relevant by both groups of citizens. For the sake of simplicity, I henceforth refer to these two policy plans as "air-pollution" and "recycling" policies.

Data from the survey experiment confirms the empirical assumption that the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents (`r filter(SVIVA2_01,AREA==1)%>%summarise(mean(RELEVANCE_air_obs))%>%max()%>%round(1)` [`r filter(SVIVA2_01,AREA==1)%>%summarise(mean(RELEVANCE_air_obs)-1.96*sd(RELEVANCE_air_obs)/sqrt(n()))%>%max()%>%round(1)`, `r filter(SVIVA2_01,AREA==1)%>%summarise(mean(RELEVANCE_air_obs)+1.96*sd(RELEVANCE_air_obs)/sqrt(n()))%>%max()%>%round(1)`] compared with `r filter(SVIVA2_01,AREA==0)%>%summarise(mean(RELEVANCE_air_obs))%>%max()%>%round(1)` [`r filter(SVIVA2_01,AREA==0)%>%summarise(mean(RELEVANCE_air_obs)-1.96*sd(RELEVANCE_air_obs)/sqrt(n()))%>%max()%>%round(1)`, `r filter(SVIVA2_01,AREA==0)%>%summarise(mean(RELEVANCE_air_obs)+1.96*sd(RELEVANCE_air_obs)/sqrt(n()))%>%max()%>%round(1)`] on a scale of 1-7). As for the second, recycling policy, it is perceived by both groups as relatively personally relevant by both groups, with no statistically significant differences between them (`r filter(SVIVA2_01,AREA==1)%>%summarise(mean(RELEVANCE_waste_obs))%>%max()%>%round(1)` versus `r filter(SVIVA2_01,AREA==0)%>%summarise(mean(RELEVANCE_waste_obs))%>%max()%>%round(1)`). Participants were also asked to describe the most salient environmental issue in their particular area of residence. Indeed, most of the Haifa-Bay subjects regarded in their responses to the air-pollution generated from the industrial area, and many of them linked it to higher morbidity rates in the area.[^footnote_environment_concern] In a prior survey, I empirically confirmed the assumption that the Haifa-bay residents do not have significantly greater knowledge about the air-pollution policy (see supplementary appendix ##TBA##).

[^footnote_environment_concern]: Subjects from center cities regarded to a variety of issues, including recycling, traffic jams and cleanliness, whereas relatively few of them mentioned air-pollution as a prominent environmental concern in their area.

<br>

##*Sample selection and recruitment*

The survey was conducted in late January 2018. To capture natural variation in the personal relevance of the air-pollution policy, I sampled citizens from the major cities in the Haifa-Bay area (*Haifa*, *Nesher*, *Qiryat-Hayiim*, *Qiryat-Bialik*, *Qiryat-Mozkin*, *Qiryat-Ata* and *Qiryat-Yam*), and matched them with citizens from cities with a similar profile and population size at the center of the country (*Natania*, *Petach-Tikva* and *Rishon-Lezion*). I will henceforth refer to these two groups of cities as Haifa-bay and center.

I conducted the survey experiment through Qualtrics Survey Software. To select and recruit residents of these specific cities, I was assisted by an Israeli internet research panel company *iPanel*. The online survey link was sent by iPanel, between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%).[^footnote_responserate] Of these, I later filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and respondents under the age of 18 (n=`r filter_age`). These filters are not significantly associated with the areas or the experimental conditions. Ultimately, following this rigorous screening, I ended up with a sizable sample of `r N_1` subjects, of whom `r filter(SVIVA2_01,AREA_reside_haifa==1)%>%nrow()` reside in the Haifa-bay cities, and `r filter(SVIVA2_01,AREA_reside_haifa==0)%>%nrow()` reside in the center cities.

[^footnote_responserate]: According to iPanel's reporting, 5,077 people were invited to fill the survey. Of these people, approximately 330 women entered the survey, but were eventually denied of access, to avoid the over-representation of women in the sample. Hence, the response rate, when reducing those who were denied of access, is slightly higher  and reaches `r round((N_0)/(5077-330)*100,1)`%.

<br>

## *Procedure*

To disguise the purpose of the experiment, the survey was presented to the participants as a study about citizens' attitudes regarding environmental policies in general, as opposed to attitudes about EPM or the specific two policies. The participants were first asked pre-manipulation questions about their political ideology and their trust in government ministries in general and evaluation of their performance. Thereafter, they were asked about their interest in environmental policy issues, the extent to which they follow these issues in the media, their preference for more/less government investment in this field. They were also asked to describe the most salient environmental issue in their particular area of residence.[^footnote_relevance_manipulation]

[^footnote_relevance_manipulation]: The latter questions about environmental policy was initially designed as a manipulation of for perceived personal relevance (specifically for Haifa-Bay subjects with regard to the air-pollution policy). The questions which are mentioned above were included in the treatment group, whereas subjects in the control group were asked instead equivalent questions about their field of occupation. However, manipulation checks suggested that these priming questions did not enhance the perceived personal relevance of these policies for the participants (either from Haifa-Bay or center cities). Therefore, I will not regard to this treatment in the paper.

Thereafter, participants were told that they will be presented with two government policy plans regarding specific environmental issues, obtained from EPM's official publication of its annual work plan for the year 2018. Each policy plan included the title of the policy goal (i.e. "reducing the levels of air pollution in the Haifa bay"), followed by short descriptions of two actions to be taken by EPM in order to fulfill this goal (30-50 words each).

The air-pollution and recycling policy plans were presented to each subject in a random order. Per each policy, the subjects were randomly allocated to one of two conditions of substantive information (hereafter: information): strong policy (i.e. logically persuasive policy plan) and weak policy (i.e. logically unpersuasive). The policy activities which were mentioned in the strong policy plans condition were taken from the real EPM work plan and from other reports and official sources of EPM. The proposed actions for the air-pollution strong policy involved increasing the supervision of polluting factories, and reducing diesel smoke emissions by vehicles in the Haifa-Bay area. The strong recycling policy regarded increasing the supervision of manufacturers and reducing the use of disposable bags. The activities that were included in the weak policy plans condition were fictional, and were designed to represent measures which are likely to be perceived by subjects as extremely incompatible with achieving the policy goals, yet still sufficiently reliable. Accordingly, the weak policy plans regarded reducing of regulatory burden on the factories/ manufacturers, and raising citizens' awareness of the efforts to improve air quality/reduce waste. The full texts of the weak and strong conditions for the two policy plans is displayed in Appendix A. For each subject, one of the two policies was strong and the other was weak, which enables me to compare both between- and within-subjects. For example, a subject who first received a strong air-pollution policy, later received the weak recycling policy, and vice versa.

The participants are also randomly assigned to one of three conditions of symbolic elements in the communication (hereafter: symbols): treatment, control no-symbols and control fake-symbols. In the symbol treatment condition, the two policies (again, one weak and one strong) are displayed with the above-mentioned three symbolic elements: the titles and the footer are colored with green, EPM logo is displayed at the top of the page, and the celebrities' images from the campaigns are attached and labeled "EPM advertisement". Each communication included the image of one celebrity: Rosenblum's image from the campaign for reducing the use of disposable bags is attached to the recycling policy, given the relevance of the campaign to the policy plan; Friedman's image from the campaign of "starting to think green", which was more general and regarded to a variety of environmental issues, was attached to the air-pollution policy. In the control no-symbols condition, the two policies are displayed in a minimal black and white design, and without the logo and the celebrities images. Finally, in the control fake-symbols condition, the green color is replaced with blue, the EPM logo is replaced by a fake logo and the images of celebrities from the campaigns are replaced with two edited images of unfamiliar people. The purpose of the fake-logo condition is to isolate the symbolic meaning of the EPM's graphic identifiers and distinguish it from the possible effect of their aesthetic design. Accordingly, I specifically designed the "fake" symbols to that they will resemble the real symbols and have equivalent aesthetic qualities, without activating emotions and associations which can be attached to the government ministry and environmental issues. In that sense, it can be regarded as a placebo. In two prior surveys, I empirically confirmed that the fake symbols are unfamiliar, do not have strong positive or negative affect, and are not strongly associated with environment policy and EPM (supplementary appendix ##TBA##). The symbols conditions are displayed in the Appendix B.

Each policy plan was followed by questions about respondents' trust in the policy. Afterwards, the subjects were asked about their perceived level of elaboration on the last policy that was presented to them, conducted a memory test. They were also asked about the extent to which they perceive the two policies as personally relevant to themselves. Thereafter, I included an instructional manipulation check (IMC) item, which aims to assess participants’ diligence. After a set of demographic questions, I conducted a manipulation check, in which I asked the participants whether they recognize the Ministry's logo, and the two images of the celebrities. Finally, the participants are debriefed about the manipulations, and the possible deception. An English translation of the full survey
is available from the supplementary appendix.

Altogether, for each policy plan, the survey experiment has a factorial design of 3 (symbols: real symbols, fake-symbols, no-symbols) x 2 (information: strong vs. weak policy). The experimental groups are fairly balanced with respect to demographics, pre-manipulation variables and the two areas. The assignment to the different experimental groups is graphically presented in Figure 3.

<br>
**Figure 3: Experimental design for policy plan communications**
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/experiment_flow.png)

##*Operationalization of variables*

```{r, echo=FALSE, warning=FALSE, message=F}
alpha_air_trust = psych::alpha(select(SVIVA2_01, TRUST_air_q1,
                               TRUST_air_q2,
                               TRUST_air_q3,
                               TRUST_air_q4,
                               TRUST_air_q5,
                               TRUST_air_q6))[["total"]][["raw_alpha"]] %>% round(2)

alpha_waste_trust = psych::alpha(select(SVIVA2_01, TRUST_waste_q1,
                               TRUST_waste_q2,
                               TRUST_waste_q3,
                               TRUST_waste_q4,
                               TRUST_waste_q5,
                               TRUST_waste_q6))[["total"]][["raw_alpha"]] %>% round(2)

cor_elaboration_air = cor.test(SVIVA2_01$MEMORY_air_score,SVIVA2_01$ELABORATION_air_time_log,method="spearman")$estimate %>% mean() %>% round(2)

cor_elaboration_waste = cor.test(SVIVA2_01$MEMORY_waste_score,SVIVA2_01$ELABORATION_waste_time_log,method="spearman")$estimate %>% mean() %>% round(2)

cor_RT_delta = cor.test(SVIVA2_01$ELABORATION_total_time_log,
                          SVIVA2_01$WITHIN_DELTA)$estimate %>% mean() %>% round(2)

cor_memory_delta = cor.test(SVIVA2_01$MEMORY_total,
                          SVIVA2_01$WITHIN_DELTA,method="spearman")$estimate %>% mean() %>% round(2)

t_air_information = t.test(TRUST_air_INDEX~INFORMATION_air,SVIVA2_01)
t_waste_information = t.test(TRUST_waste_INDEX~INFORMATION_waste,SVIVA2_01)

```

My outcome variable is citizens' trust in policy (or the perceived trustworthiness of the policy). This variable is measured in the survey for each policy plan using a composite index of six likert scale items (Cronbach’s alpha = `r alpha_air_trust` and `r alpha_waste_trust` for the air-pollution and waste policies respectively), which were adopted and slightly modified from validated trust scales used by previous public administration studies on trust in government [@grimmelikhuijsen_2013; @grimmelikhuijsen_2014; @karens_2016; @porumbescu_2016; @grimmelikhuijsen_2017]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7(strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan.

To assess the levels of thinking or elaboration on the policy plans, I used two measures: First, I measure the number of seconds participants spent on scanning each of the policy plans (hereafter – reaction time or RT), log transformed to reduce skewness. Second, I used a memory test, asking participants whether a list of four items were included in the policy plan, of which two were/were not included (participants could also select "I don't know"). The memory score is calculated as the number of correct items selected by the participant minus the number of incorrect items, with a minimum score of zero. Accordingly, participants could receive a score of 0, 1 or 2.[^footnote_elaboration] The two measures of elaboration are correlated with each other (Spearman rho = `r cor_elaboration_air` and `r cor_elaboration_waste` for the air-pollution and recycling policies respectively), which substantiates their validity. Higher reaction time and memory scores were also positively associated with lower delta (within-subjects) between trust in strong and weak policy plans, which indicates the extent to which subjects payed attention to the logical quality of the policy plans (correlations with reaction time and memory score, respectively, are r = `r cor_RT_delta` and rho = `r cor_memory_delta`).

[^footnote_elaboration]: In addition, I included in the survey a battery of five items that measure participants' subjective perception of their level of elaboration. I initially sought to use these items as an additional measure for elaboration. However, I decided to forgo this subjective measure, due to its lack of correlation with the reaction time and memory test score, and its possible bias by people's disinclination to admit that they did not thoroughly read the policy plans.

My main independent variables are the above mentioned manipulations for information and symbols, and the distinction between the Haifa-Bay and center, as a proxy for the extent to which they perceive the air-pollution policy as personally relevant. The latter was coded according to respondents’ reporting on residence, work or studying in one of the Haifa-Bay cities.

<br>

# Results

**Figure 4** graphically shows the distributions of participants' trust in the air-pollution and recycling policy plans. Both distributions are left skewed, with the majority of subjects reporting 4 or more (on a 1-7 scale). Yet, participants were more likely to report low levels of trust in the air-pollution policy, and accordingly its mean is slightly lower than the recycling policy (`r SVIVA2_01%>%summarise(mean(TRUST_air_INDEX))%>%max()%>%round(2)` [`r SVIVA2_01%>%summarise(mean(TRUST_air_INDEX)-1.96*sd(TRUST_air_INDEX)/sqrt(n()))%>%max()%>%round(2)`, `r SVIVA2_01%>%summarise(mean(TRUST_air_INDEX)+1.96*sd(TRUST_air_INDEX)/sqrt(n()))%>%max()%>%round(2)`], compared with `r SVIVA2_01%>%summarise(mean(TRUST_waste_INDEX))%>%max()%>%round(2)` [`r SVIVA2_01%>%summarise(mean(TRUST_waste_INDEX)-1.96*sd(TRUST_waste_INDEX)/sqrt(n()))%>%max()%>%round(2)`, `r SVIVA2_01%>%summarise(mean(TRUST_waste_INDEX)+1.96*sd(TRUST_waste_INDEX)/sqrt(n()))%>%max()%>%round(2)`]). Higher levels of trust are associated with high trust vis a vis government and left-wing ideology.

**Figure 4: Distribution of trust in the two policies**

```{r,echo=FALSE,warning=FALSE,message=FALSE, out.width=c('50%', '50%'), fig.show='hold'}
ggplot(SVIVA2_01, aes(x=TRUST_air_INDEX)) + 
  geom_histogram(aes(y = (..count..)/sum(..count..)),binwidth = 1,color="black")+
  scale_x_continuous(name = "Trust in air-pollution policy",breaks = c(1:7),labels = c("low","2","3","4","5","6","high"))+
  scale_y_continuous(name = "Percentage",limits = c(0,0.4))+
  guides(fill=FALSE)+
  ggtitle("") +
  theme_classic()

ggplot(SVIVA2_01, aes(x=TRUST_waste_INDEX)) + 
  geom_histogram(aes(y = (..count..)/sum(..count..)),binwidth = 1,color="black")+
  scale_x_continuous(name = "Trust in recycling policy",breaks = c(1:7),labels = c("low","2","3","4","5","6","high"))+
  scale_y_continuous(name = "",limits = c(0,0.4))+
  guides(fill=FALSE)+
  ggtitle("") +
  theme_classic()
```

I now turn to examine my theoretical hypotheses. To reiterate, my first two hypotheses regard the role of symbolic elements as peripheral cues that evoke positive emotions, and thereby increase citizens' trust (H1), mainly among those who do not perceive the communication as personally relevant (H2). Thereafter, I theorize that symbols can increase trust, indirectly, by decreasing the amount of thinking and thus compensating for logically unpersuasive communications. Accordingly, hypothesis 3 focuses on the interaction between symbols and information, and hypothesis 4 focuses on its mediation by lower levels of elaboration.

```{r, echo=FALSE,warning=FALSE,message=FALSE}

x_realsymbol_air = filter(SVIVA2_01,SYMBOL==2)%>%summarise(mean(TRUST_air_INDEX)) %>% max()
x_fakesymbol_air = filter(SVIVA2_01,SYMBOL==1)%>%summarise(mean(TRUST_air_INDEX)) %>% max()
x_nosymbol_air = filter(SVIVA2_01,SYMBOL==0)%>%summarise(mean(TRUST_air_INDEX)) %>% max()
x_realsymbol_waste = filter(SVIVA2_01,SYMBOL==2)%>%summarise(mean(TRUST_waste_INDEX)) %>% max()
x_fakesymbol_waste = filter(SVIVA2_01,SYMBOL==1)%>%summarise(mean(TRUST_waste_INDEX)) %>% max()
x_nosymbol_waste = filter(SVIVA2_01,SYMBOL==0)%>%summarise(mean(TRUST_waste_INDEX)) %>% max()

t_air_symbol_nosymbol = t.test(TRUST_air_INDEX~SYMBOL,filter(SVIVA2_01,SYMBOL!=1))
t_air_symbol_fakesymbol = t.test(TRUST_air_INDEX~SYMBOL,filter(SVIVA2_01,SYMBOL!=0))
t_waste_symbol_nosymbol = t.test(TRUST_waste_INDEX~SYMBOL,filter(SVIVA2_01,SYMBOL!=1))
t_waste_symbol_fakesymbol = t.test(TRUST_waste_INDEX~SYMBOL,filter(SVIVA2_01,SYMBOL!=0))

d_nosymbol_air = ((x_realsymbol_air-x_nosymbol_air)/(filter(SVIVA2_01,SYMBOL!=1)%>%summarise(sd(TRUST_air_INDEX))%>%max())) %>% round(2)
d_nosymbol_waste = ((x_realsymbol_waste-x_nosymbol_waste)/(filter(SVIVA2_01,SYMBOL!=1)%>%summarise(sd(TRUST_waste_INDEX))%>%max())) %>% round(2)

t_symbol_fakesymbol = t.test(trust~SYMBOL,filter(SVIVA2_01_comb,SYMBOL!=0))

```

## *The effect of symbols on trust in policy*

Beginning with my first hypothesis, I expect that subjects will report higher levels of trust in the policy plans, when they are presented with the real symbols, compared with the two control groups. **Figure 5** shows the descriptive comparison of trust between these experimental groups. The black points and error bars represent the mean and 95% confidence intervals, and the box-plots show the distribution (position of quartiles and outliers). In both policies, subjects in the real symbols condition reported higher levels of trust compared with the no-symbols condition (differences are `r (x_realsymbol_air-x_nosymbol_air)%>%round(2)` [`r t_air_symbol_nosymbol[["conf.int"]][1] %>% round(2)`,`r t_air_symbol_nosymbol[["conf.int"]][2] %>% round(2)`] and `r (x_realsymbol_waste-x_nosymbol_waste)%>%round(2)` [`r t_waste_symbol_nosymbol[["conf.int"]][1] %>% round(2)`,`r t_waste_symbol_nosymbol[["conf.int"]][2] %>% round(2)`] for the air-pollution and recycling policies, respectively). These differences are equivalent to `r d_nosymbol_air` and `r d_nosymbol_waste` standard deviations. The fake symbols also increased participants' trust in the policy plans, but to a lesser extent. Subjects in the real symbols reported higher trust compared with the fake symbols, but the differences are marginally significant (*p* = `r (t_air_symbol_fakesymbol[["p.value"]]/2)%>%round(2)`-`r (t_waste_symbol_fakesymbol[["p.value"]]/2)%>%round(2)`). Below, I also combine the trust in the two policies (i.e. 2 observations per each subject). In these multilevel analyses, the effect of the real symbols compared with the fake symbols is significant at the 90% level.

**Figure 5: The effect of symbolic elements on trust in policy**

```{r, echo=FALSE,warning=FALSE,message=FALSE, fig.show='hold', fig.width=6,fig.height=4}
SVIVA2_01_comb = SVIVA2_01_comb %>% 
  mutate(POLICY_t = Recode(policy,"'TRUST_air_INDEX'='Air-pollution';'TRUST_waste_INDEX'='Recycling'"))

ggplot(SVIVA2_01_comb, aes(factor(SYMBOL_n),trust))+ 
  theme_classic()+
  geom_boxplot(width=0.25,color = "gray70")+
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0.1,size=0.5) + 
  stat_summary(fun.y = mean, geom = "point", size=1.5)+
  stat_summary(fun.y = mean, aes(group = 1), geom = "line", size=0.5)+
  facet_grid(~factor(POLICY_t))+
  scale_x_discrete(name="",labels=c("No symbols","Fake symbols","Real symbols"))+
  scale_y_continuous(name = "Trust in EPM policy",breaks = seq(1,7,by=0.5))

```

```{r, echo=FALSE,warning=FALSE,message=FALSE}

q_diff = c(
  ((quantile(SVIVA2_01$TRUST_air_INDEX[SVIVA2_01$SYMBOL==2])[3]- quantile(SVIVA2_01$TRUST_air_INDEX[SVIVA2_01$SYMBOL==0])[3])%>%max()),

((quantile(SVIVA2_01$TRUST_waste_INDEX[SVIVA2_01$SYMBOL==2])[3]- quantile(SVIVA2_01$TRUST_waste_INDEX[SVIVA2_01$SYMBOL==0])[3])%>%max()),

((quantile(SVIVA2_01$TRUST_air_INDEX[SVIVA2_01$SYMBOL==2])[4]- quantile(SVIVA2_01$TRUST_air_INDEX[SVIVA2_01$SYMBOL==0])[4])%>%max()),

((quantile(SVIVA2_01$TRUST_waste_INDEX[SVIVA2_01$SYMBOL==2])[4]- quantile(SVIVA2_01$TRUST_waste_INDEX[SVIVA2_01$SYMBOL==0])[4])%>%max()))

```

Overall, the results are largely consistent with my hypothesis that familiar symbolic elements can increase trust by evoking positive associations that are transferred to government organizations and policies. The graphic elements in the fake symbols condition also increased participants' trust through the peripheral route, but to a lesser extent, presumably due to their unfamiliarity, and their lack of strong positive affect.

**Figure 5** also shows that the distribution of trust in policy in the real symbols group is more concentrated. The comparison of the quartiles in the different groups indicates that the increase in the average was mainly due to the increase in the first quartile (i.e. the lower 25% observations). Compared with the no symbols, the first quartile increased by `r (quantile(SVIVA2_01$TRUST_air_INDEX[SVIVA2_01$SYMBOL==2])[2]- quantile(SVIVA2_01$TRUST_air_INDEX[SVIVA2_01$SYMBOL==0])[2])%>%max()%>%round(1)` and `r (quantile(SVIVA2_01$TRUST_waste_INDEX[SVIVA2_01$SYMBOL==2])[2]- quantile(SVIVA2_01$TRUST_waste_INDEX[SVIVA2_01$SYMBOL==0])[2])%>%max()%>%round(1)`, with respect to the air-pollution and recycling policies, respectively. By comparison, the medians and 3rd quartiles increased only by `r q_diff %>% min() %>% round(1)`-`r q_diff %>% max() %>% round(1)`. On the face of it, this comparison suggests that symbolic elements are more effective in shifting people from low to medium levels of trust, than shifting them from medium to high levels. Stated otherwise, symbols may be more effective in reducing distrust, as opposed to increasing trust. If this explanation holds, it implies that symbols have a decreasing marginal effect, i.e., they are generally more effective when trust is low in the first place.

<br>

## *The interaction bewteen symbols and personal relevance*

According to my second hypothesis, I expect that symbols' effect on trust in air-pollution policy will be stronger among the center subjects, insofar as they do not perceive this issue personally relevant. Conversely, regarding the recycling policy, the effect of the symbols is expected to be similar for residents of both areas.

Before analyzing the interaction, I explore the differences in the distribution of trust across the two areas. In general, participants from Haifa-Bay were more likely to report low levels of trust in the air-pollution policy (*MEAN* = `r filter(SVIVA2_01,AREA==1)%>%summarise(mean(TRUST_air_INDEX))%>%max()%>%round(2)` [`r filter(SVIVA2_01,AREA==1)%>%summarise(mean(TRUST_air_INDEX)-1.96*sd(TRUST_air_INDEX)/sqrt(n()))%>%max()%>%round(2)`, `r filter(SVIVA2_01,AREA==1)%>%summarise(mean(TRUST_air_INDEX)+1.96*sd(TRUST_air_INDEX)/sqrt(n()))%>%max()%>%round(2)`], *Median* = `r filter(SVIVA2_01,AREA==1)%>%summarise(median(TRUST_air_INDEX))%>%max()%>%round(2)`), compared with participants from the center (*MEAN* = `r filter(SVIVA2_01,AREA==0)%>%summarise(mean(TRUST_air_INDEX))%>%max()%>%round(2)` [`r filter(SVIVA2_01,AREA==0)%>%summarise(mean(TRUST_air_INDEX)-1.96*sd(TRUST_air_INDEX)/sqrt(n()))%>%max()%>%round(2)`, `r filter(SVIVA2_01,AREA==0)%>%summarise(mean(TRUST_air_INDEX)+1.96*sd(TRUST_air_INDEX)/sqrt(n()))%>%max()%>%round(2)`], *Median* = `r filter(SVIVA2_01,AREA==0)%>%summarise(median(TRUST_air_INDEX))%>%max()%>%round(2)`). The distribution among the Haifa-Bay subjects is also more dispersed (*SD* = `r filter(SVIVA2_01,AREA==1)%>%summarise(sd(TRUST_air_INDEX))%>%max()%>%round(2)` vs. `r filter(SVIVA2_01,AREA==0)%>%summarise(sd(TRUST_air_INDEX))%>%max()%>%round(2)`). The distribution of trust in the recycling policy is fairly similar in both areas, with slightly higher trust among center subjects (*Mean* = `r filter(SVIVA2_01,AREA==0)%>%summarise(mean(TRUST_waste_INDEX))%>%max()%>%round(2)` vs. `r filter(SVIVA2_01,AREA==1)%>%summarise(mean(TRUST_waste_INDEX))%>%max()%>%round(2)`, *p*<.1). These differences may suggest that Haifa-Bay residents tend to have prior negative attitudes towards EPM's actions regarding the air-pollution in their area, but not towards EPM or environmental policy in general.

In **Table 1**, I examine the interaction between symbols and the two areas using OLS regression analyses. With regard to each of the two policy plans, I regress the trust index on the symbols manipulation and the allocation to areas (Models 1.1 and 1.3), and thereafter I add the interaction between them (models 1.2 and 1.4).

**Table 1: Regression analyses**

```{r,results="asis",echo=FALSE,warning=FALSE,message=FALSE}

mod_air_1.1 = lm(TRUST_air_INDEX~SYMBOL_t,data=SVIVA2_01_comb)
mod_air_1.2 = update(mod_air_1.1,. ~ .+AREA_center)
mod_air_1.3 = update(mod_air_1.1,. ~ .+AREA_center*SYMBOL_t)

mod_waste_1.1 = lm(TRUST_waste_INDEX~SYMBOL_t,data=SVIVA2_01_comb)
mod_waste_1.2 = update(mod_waste_1.1,. ~ .+AREA_center)
mod_waste_1.3 = update(mod_waste_1.1,. ~ .+AREA_center*SYMBOL_t)

stargazer(mod_air_1.2,mod_air_1.3,
          mod_waste_1.2,mod_waste_1.3,
          type="html",style = "apsr",
          dep.var.labels=c("Trust in air-pollution policy","Trust in recycling policy"),
          covariate.labels=c("Real symbols",
                             "Fake-symbols",
                             "Area (0=Haifa-Bay; 1=center)",
                             "Area x Real symbols",
                             "Area x Fake symbols"),
          column.labels = c("(1.1)", "(1.2)","(1.3)","(1.4)"),
          model.numbers=F,
          omit.stat = c("rsq","ser", "f"),
          initial.zero = FALSE,
          notes = "*Notes*: Table entries are nonstandardized OLS-regression coefficients, with Standard errors in parentheses. The reference category for the symbols manipulation conditions is the control (no symbols).",
          notes.append = FALSE)

mod_air_1.1.r = lm(TRUST_air_INDEX~SYMBOL_t.r,data=SVIVA2_01_comb)
mod_air_1.2.r = update(mod_air_1.1.r,. ~ .+AREA_center)
mod_air_1.3.r = update(mod_air_1.1.r,. ~ .+AREA_center*SYMBOL_t.r)

mod_waste_1.1.r = lm(TRUST_waste_INDEX~SYMBOL_t.r,data=SVIVA2_01_comb)
mod_waste_1.2.r = update(mod_waste_1.1.r,. ~ .+AREA_center)
mod_waste_1.3.r = update(mod_waste_1.1.r,. ~ .+AREA_center*SYMBOL_t.r)
```

&ast;*p*<.1; &ast;&ast;*p*<.05; &ast;&ast;&ast;*p*<.01

These analyses suggest that the effect of the real symbols on the air-pollution policy was not weaker among the Haifa-bay subjects. If anything, the negative coefficient of the interaction term *Area x Real symbols* in Model 1.2 (*Coef* = `r mod_air_1.3[["coefficients"]][["SYMBOL_t1:AREA_center"]]%>%round(2)`, *p*<.1) indicates that the symbols were perhaps more effective on those who reside in the Haifa-bay. Furthermore, Model 1.4 also shows that the effect was greater among Haifa-Bay subjects for the recycling policy, which suggests that the differences between the areas is not explained by their varying levels of personal relevance. Rather, these differences could be linked to the predispositions of Haifa-Bay subjects to have lower levels of trust in the first place.[^footnote_higher_effect_haifa]

In sum, the experimental results suggest that symbolic elements in communications increase citizens' trust in the policies presented in the communications, and that this effect is not moderated by the perceived personal relevance of the issue. In order to further explore the robustness of my null findings with regard to the interaction with personal relevance, I conducted additional analyses (see supplementary appendix). First, in addition to the distinction between the areas, I supplemented the above operationalization of perceived personal relevance with subjects' reporting of the perceived personal relevance of these policy plans. Second, I added a set of individual-level control variables, and other factors related to interest in environmental policy in general. Third, I further explored whether Haifa-bay subjects were more likely to scrutinize the air-pollution policy by testing the interaction between the areas and the information manipulation, and by comparing the levels of elaboration (reaction time and memory-test score) across the two areas, and finally the delta (within-subjects) between trust in strong and weak policy plans. All these analyses yielded null findings, which suggest that those who perceive these policies issues as more personally relevant were not more likely to elaborate on them, and were not less likely to be influenced by the symbolic elements.

[^footnote_higher_effect_haifa]: This tentative explanation is supported by the observed decreased marginal reduction in the percentage of subjects who reported very low levels of trust in the policies. For example, in the Haifa-Bay, the percentage of those who reported a trust of less than 2 in the air-pollution policy decreased from `r ((filter(SVIVA2_01_haifa,SYMBOL==0,TRUST_air_INDEX<2) %>% nrow())/(filter(SVIVA2_01_haifa,SYMBOL==0)%>%nrow())*100)%>%round(1)`% in the no symbols group to `r ((filter(SVIVA2_01_haifa,SYMBOL==2,TRUST_air_INDEX<2) %>% nrow())/(filter(SVIVA2_01_haifa,SYMBOL==2)%>%nrow())*100)%>%round(1)`%, and in the center, it decreased from `r ((filter(SVIVA2_01_center,SYMBOL==0,TRUST_air_INDEX<2) %>% nrow())/(filter(SVIVA2_01_center,SYMBOL==0)%>%nrow())*100)%>%round(1)`% to `r ((filter(SVIVA2_01_center,SYMBOL==2,TRUST_air_INDEX<2) %>% nrow())/(filter(SVIVA2_01_center,SYMBOL==2)%>%nrow())*100)%>%round(1)`%. With regard to the recycling policy, the percentage decreased from `r ((filter(SVIVA2_01_haifa,SYMBOL==0,TRUST_waste_INDEX<2) %>% nrow())/(filter(SVIVA2_01_haifa,SYMBOL==0)%>%nrow())*100)%>%round(1)`% to `r ((filter(SVIVA2_01_haifa,SYMBOL==2,TRUST_waste_INDEX<2) %>% nrow())/(filter(SVIVA2_01_haifa,SYMBOL==2)%>%nrow())*100)%>%round(1)`% and in the center it decreased from `r ((filter(SVIVA2_01_center,SYMBOL==0,TRUST_waste_INDEX<2) %>% nrow())/(filter(SVIVA2_01_center,SYMBOL==0)%>%nrow())*100)%>%round(1)`% to `r ((filter(SVIVA2_01_center,SYMBOL==2,TRUST_waste_INDEX<2) %>% nrow())/(filter(SVIVA2_01_center,SYMBOL==2)%>%nrow())*100)%>%round(1)`%.

```{r, echo=FALSE,warning=FALSE,message=FALSE}
x_realsymbol = filter(SVIVA2_01_comb,SYMBOL==2)%>%summarise(mean(trust)) %>% max()
x_fakesymbol = filter(SVIVA2_01_comb,SYMBOL==1)%>%summarise(mean(trust)) %>% max()
x_nosymbol = filter(SVIVA2_01_comb,SYMBOL==0)%>%summarise(mean(trust)) %>% max()
sd_trust = SVIVA2_01_comb%>%summarise(sd(trust))%>%max()

x_realsymbol_delta = filter(SVIVA2_01,SYMBOL==2)%>%summarise(mean(WITHIN_DELTA)) %>% max()
x_fakesymbol_delta = filter(SVIVA2_01,SYMBOL==1)%>%summarise(mean(WITHIN_DELTA)) %>% max()
x_nosymbol_delta = filter(SVIVA2_01,SYMBOL==0)%>%summarise(mean(WITHIN_DELTA)) %>% max()
sd_trust_delta = SVIVA2_01%>%summarise(sd(WITHIN_DELTA))%>%max()

```

Given the lack of interaction between symbols and personal relevance, and in order to simplify the analyses of the remaining hypotheses, I continue the rest of my analyses while combining the two policies and the two areas together. Accordingly, I transformed the dataset into a multilevel structure of policies nested within subjects (N observations = `r SVIVA2_01_comb %>% nrow()`, N subjects = `r SVIVA2_01_comb%>%distinct(IP)%>%nrow()`).[^footnote_combined_analyses]

[^footnote_combined_analyses]: In the supplementary appendix I further replicate the analyses below, while separating between the two policies.

## *The interaction bewteen symbols and information*

I now turn to test hypothesis 3, which regards the interaction between symbols and the substantive information in the communication. Empirically, I expect that the effect of the real symbols will be greater for the weak policy, compared with the strong policy. In **Table 2**, I examine this hypothesis via regression analyses. To account for the multilevel structure of the data, I use GLS regressions with random intercept at the subject level (ICC1 = `r multilevel::ICC1(aov(trust~as.factor(IP),data=SVIVA2_01_comb))%>%round(2)`, ICC2 = `r multilevel::ICC2(aov(trust~as.factor(IP),data=SVIVA2_01_comb))%>%round(2)`). In Model 2.1 I regress the policy trust index on the symbols and information manipulations. Thereafter, in Model 2.2, I add the interaction between them. In Model 3.3, I also control for the different effects of the information manipulation on the air-pollution and recycling policies.

**Table 2: Regression analyses**

```{r,results="asis",echo=FALSE,warning=FALSE,message=FALSE}
library(lme4)

mod_2.1 = lmer(trust~SYMBOL_t+INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)
mod_2.2 = lmer(trust~SYMBOL_t*INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)
mod_2.3 = lmer(trust~SYMBOL_t*INFORMATION_weak+policy*INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)

stargazer(mod_2.1,mod_2.2,mod_2.3,
          type="html",style = "apsr",
          order = c(1,2,3,5,6,4,7),
          dep.var.labels=c("Trust in policy (combined)"),
          covariate.labels=c("Real symbols",
                             "Fake-symbols",
                             "Weak policy (0=strong policy)",
                             "Weak policy x Real symbols",
                             "Weak policy x Fake symbols",
                             "Recycling policy (0=air-pollution)",
                             "Weak policy x Recycling policy"),
          #omit.stat = c("aic","bic"),
          column.labels = c("(2.1)", "(2.2)","(2.3)"),
          initial.zero = FALSE,
          notes = "*Notes*: Table entries are non-standardized random-effect GLS regression coefficients, clustered at the subject level. Standard errors are in parentheses. The reference category for the symbols manipulation conditions is the control (no symbols).",
          notes.append = FALSE)

mod_2.2.r_1 = lmer(trust~SYMBOL_t.r*INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)
mod_2.3.r_1 = lmer(trust~SYMBOL_t.r*INFORMATION_weak+policy*INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)

mod_2.2.r_2 = lmer(trust~SYMBOL_t*INFORMATION+(1|IP),data=SVIVA2_01_comb)
mod_2.3.r_2 = lmer(trust~SYMBOL_t*INFORMATION+policy*INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)

mod_2.2.r_3 = lmer(trust~SYMBOL_t.r*INFORMATION+(1|IP),data=SVIVA2_01_comb)
mod_2.3.r_3 = lmer(trust~SYMBOL_t.r*INFORMATION+policy*INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)

mod_2.2.r_4 = lmer(trust~SYMBOL_t.r_2*INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)
mod_2.3.r_4 = lmer(trust~SYMBOL_t.r_2*INFORMATION_weak+policy*INFORMATION_weak+(1|IP),data=SVIVA2_01_comb)
```

&ast;*p*<.1; &ast;&ast;*p*<.05; &ast;&ast;&ast;*p*<.01

The main effect of the weak policy in Model 2.1 is negative and significant, (*Coef* = `r mod_2.1%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.1%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())-1.96*(mod_2.1%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.1%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())+1.96*(mod_2.1%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`]). The interaction between weak policy and real symbols is positive and significant in Model 2.2 (*Coef* = `r mod_2.2%>%tidy()%>%filter(term=="SYMBOL_t1:INFORMATION_weak")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.2%>%tidy()%>%filter(term=="SYMBOL_t1:INFORMATION_weak")%>%select(estimate)%>%max())-1.96*(mod_2.2%>%tidy()%>%filter(term=="SYMBOL_t1:INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.2%>%tidy()%>%filter(term=="SYMBOL_t1:INFORMATION_weak")%>%select(estimate)%>%max())+1.96*(mod_2.2%>%tidy()%>%filter(term=="SYMBOL_t1:INFORMATION_weak")%>%select(std.error)%>%max()))%>%round(2)`]), whereas the interaction with the fake symbols is insignificant. The interpretation of this model suggests that the real symbols had a greater effect on trust when the policy plan was weak (*Coef* = `r mod_2.2.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.2.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())-1.96*(mod_2.2.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.2.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())+1.96*(mod_2.2.r_2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(std.error)%>%max()))%>%round(2)`] and `r mod_2.2.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max()%>%round(2)` [`r ((mod_2.2.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())-1.96*(mod_2.2.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`, `r ((mod_2.2.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(estimate)%>%max())+1.96*(mod_2.2.r_3%>%tidy()%>%filter(term=="SYMBOL_t.r1")%>%select(std.error)%>%max()))%>%round(2)`] Compared with the no symbols and fake symbols, respectively). When the symbols were added to the strong policy plan, they had a weaker effect, which was significant only compared with the no symbols condition (*Coef* = `r mod_2.2%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max()%>%round(2)`, *p*<.1). Congruently, the symbols decreased the differences between the strong and weak policies by `r ((mod_2.2.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())/
(mod_2.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())*100) %>% round(0)`%, compared with the no symbols group (from `r mod_2.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_2.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_2.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_2.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_2.2%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`] to `r mod_2.2.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max()%>%round(2)` [`r ((mod_2.2.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())-1.96*(mod_2.2.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`, `r ((mod_2.2.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%abs()%>%max())+1.96*(mod_2.2.r_4%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(std.error)%>%abs()%>%max()))%>%round(2)`]). The interaction between symbols and information, in Model 2.2, significantly contributes to the model fit (Chi2 = `r anova(mod_2.2,mod_2.1)%>%tidy()%>%filter(term=="mod_2.2")%>%select(statistic)%>%max()%>%round(2)`, *p*<.05). To facilitate the interpretation of this interaction, **Figure 6** provides a descriptive demonstration of the effects of the real symbols on trust in policy, under weak and strong information quality. The coefficients of the manipulations and their interaction remains intact when controlling for the different policies, in Model 2.3, which suggests that in both policies the symbols had a greater effect when added to the a weak policy plan.[^footnote_interaction_symbols_policies]

[^footnote_interaction_symbols_policies]: In additional models I also added the interaction between the symbols and the policies, and the triple interaction between symbols, policies and information. These interactions are insignificant.

**Figure 6: The effect of symbolic elements on trust under weak vs. strong information**

```{r, echo=FALSE,warning=FALSE,message=FALSE, fig.width=6,fig.height=4}
ggplot(SVIVA2_01_comb, aes(factor(SYMBOL_n),trust,shape=factor(INFORMATION_n)))+ 
  theme_classic()+
  #geom_boxplot(width=0.25,color="gray70")+
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0.1,size=0.5,position=position_dodge(0.1),color="gray70") + 
  stat_summary(fun.y = mean, geom = "point", size=1.5,position=position_dodge(0.1))+
  stat_summary(fun.y = mean, geom = "line", size=0.5, aes(group = INFORMATION_n,linetype=INFORMATION_n),position=position_dodge(0.1))+
  scale_shape_manual(guide=F,values=c(19,1))+
  scale_linetype_manual(name="Information",values=c(1,4),labels=c("Weak policy","Strong policy"))+
  scale_x_discrete(name="",labels=c("No symbols","Fake symbols","Real symbols"))+
  scale_y_continuous(name = "Trust in policy",breaks = seq(1,7,by=0.1))
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}

x_realsymbol_delta=filter(SVIVA2_01,SYMBOL==2)%>%summarise(mean(WITHIN_DELTA))%>%max()
x_fakesymbol_delta=filter(SVIVA2_01,SYMBOL==1)%>%summarise(mean(WITHIN_DELTA))%>%max()
x_nosymbol_delta=filter(SVIVA2_01,SYMBOL==0)%>%summarise(mean(WITHIN_DELTA))%>%max()

t_delta_symbol_nosymbol = t.test(WITHIN_DELTA~SYMBOL,filter(SVIVA2_01,SYMBOL!=1))
t_delta_symbol_fakesymbol = t.test(WITHIN_DELTA~SYMBOL,filter(SVIVA2_01,SYMBOL!=0))
t_delta_nosymbol_fakesymbol = t.test(WITHIN_DELTA~SYMBOL,filter(SVIVA2_01,SYMBOL!=2))

```

The above analyses are based on *between-subject* comparisons for the same policy plan (i.e. strong vs. weak air-pollution policy, and strong vs. weak recycling policy). However, I can also utilize the fact that each subject received two policy plans (one strong and one weak) to assess the interaction between symbols and information, by analyzing the effect of symbols on the differences *within-subjects* - between their trust in the two policy plans. This analysis is graphically displayed in **Figure 6**. The averaged delta in the real symbols is significantly lower than in the no symbols and fake symbols control groups (differences are `r (x_realsymbol_delta-x_nosymbol_delta)%>%round(2)` [`r t_delta_symbol_nosymbol[["conf.int"]][1] %>% round(2)`,`r t_delta_symbol_nosymbol[["conf.int"]][2] %>% round(2)`] and `r (x_realsymbol_delta-x_fakesymbol_delta)%>%round(2)` [`r t_delta_symbol_fakesymbol[["conf.int"]][1] %>% round(2)`,`r t_delta_symbol_fakesymbol[["conf.int"]][2] %>% round(2)`], respectively). The differences between the two control groups are insignificant.

**Figure 6: The effect of symbolic elements on within-subject delta between strong and weak policies**

```{r, echo=FALSE,warning=FALSE,message=FALSE, fig.width=4,fig.height=4}
ggplot(SVIVA2_01, aes(factor(SYMBOL_n), WITHIN_DELTA))+
  theme_classic()+
  #geom_boxplot(width=0.25,color = "gray70")+
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0.1,size=0.5,color="#9999CC",alpha=0.7) + 
  stat_summary(fun.y = mean, geom = "point", size=2)+
  stat_summary(fun.y = mean, aes(group = 1), geom = "line", size=0.5)+
  scale_x_discrete(name="Symbolic elements",labels=c("no symbols","fake symbols","real symbols"))+
  scale_y_continuous(name = "Within subject delta", breaks = seq(0,1,by=0.1))

```

In sum, my between- and within-subject comparisons largely support my third hypothesis, and suggest that symbolic elements increase citizens' trust, in part, by attenuating the negative effect of weak or logically unpersuasive information in communications.

## *The mediating role of elaboration*

Finally, I examine my fourth hypothesis, that the real symbols increase trust in weak policy plans by decreasing the amount of thinking on the communication. Empirically, based on this hypothesis, I first expect to find a negative effect of the real symbols on my two indicators of elaboration (reaction time and memory score). Thereafter, this effect is expected to mediate the effect of symbols of trust under the weak policy, but not under the strong policy (i.e. a second stage moderated mediation model, also see Figure 1).

```{r, echo=FALSE,warning=FALSE,message=FALSE}
x_realsymbol_RT=filter(SVIVA2_01_comb,SYMBOL==2)%>%summarise(mean(ELABORATION_time))%>%max()
x_fakesymbol_RT=filter(SVIVA2_01_comb,SYMBOL==1)%>%summarise(mean(ELABORATION_time))%>%max()
x_nosymbol_RT=filter(SVIVA2_01_comb,SYMBOL==0)%>%summarise(mean(ELABORATION_time))%>%max()

t_RT_symbol_nosymbol = t.test(ELABORATION_time~SYMBOL,filter(SVIVA2_01_comb,SYMBOL!=1))
t_RT_symbol_fakesymbol = t.test(ELABORATION_time~SYMBOL,filter(SVIVA2_01_comb,SYMBOL!=0))
t_RT_nosymbol_fakesymbol = t.test(ELABORATION_time~SYMBOL,filter(SVIVA2_01_comb,SYMBOL!=2))

x_realsymbol_RT_log=filter(SVIVA2_01_comb,SYMBOL==2)%>%summarise(mean(ELABORATION_time_log))%>%max()
x_fakesymbol_RT_log=filter(SVIVA2_01_comb,SYMBOL==1)%>%summarise(mean(ELABORATION_time_log))%>%max()
x_nosymbol_RT_log=filter(SVIVA2_01_comb,SYMBOL==0)%>%summarise(mean(ELABORATION_time_log))%>%max()

cor_memory_nosymbol = cor.test(as.numeric(filter(SVIVA2_01_comb,SYMBOL!=1)$SYMBOL),
                               filter(SVIVA2_01_comb,SYMBOL!=1)$MEMORY_score, 
                               method="spearman")$estimate %>% mean() %>% round(2)

cor_memory_fakesymbol = cor.test(as.numeric(filter(SVIVA2_01_comb,SYMBOL!=0)$SYMBOL),
                               filter(SVIVA2_01_comb,SYMBOL!=0)$MEMORY_score, 
                               method="spearman")$estimate %>% mean() %>% round(2)
```

The effects of the real symbols on the elaboration measures are graphically displayed in **Figure 7**. The left side shows its effect on reaction time (logged), and the right side shows its effect on the memory-test scores. Subjects in the real symbols group spent `r (x_realsymbol_RT-x_nosymbol_RT)%>%abs()%>%round(1)`-`r (x_realsymbol_RT-x_fakesymbol_RT)%>%abs()%>%round(1)` second less on the pages of the policy plans in the online survey, compared with the two control groups.[^footnote_RT_symbols] With regard to the memory test, subjects in the real symbols group were less likely to select correct items which were included in the policy plans. `r (((filter(SVIVA2_01_comb,SYMBOL==2,MEMORY_score==0)%>%nrow())/(filter(SVIVA2_01_comb,SYMBOL==2,MEMORY_score>=0)%>%nrow()))*100)%>%round(0)`% of them received the minimum score of 0, compared with `r (((filter(SVIVA2_01_comb,SYMBOL==0,MEMORY_score==0)%>%nrow())/(filter(SVIVA2_01_comb,SYMBOL==0,MEMORY_score>=0)%>%nrow()))*100)%>%round(0)`% and `r (((filter(SVIVA2_01_comb,SYMBOL==1,MEMORY_score==0)%>%nrow())/(filter(SVIVA2_01_comb,SYMBOL==1,MEMORY_score>=0)%>%nrow()))*100)%>%round(0)`% in the no symbols and fake symbols control groups. The percentage of those who managed to correctly identify the two items dropped from `r (((filter(SVIVA2_01_comb,SYMBOL==0,MEMORY_score==2)%>%nrow())/(filter(SVIVA2_01_comb,SYMBOL==0,MEMORY_score>=0)%>%nrow()))*100)%>%round(0)`% and `r (((filter(SVIVA2_01_comb,SYMBOL==1,MEMORY_score==2)%>%nrow())/(filter(SVIVA2_01_comb,SYMBOL==1,MEMORY_score>=0)%>%nrow()))*100)%>%round(0)`% in the control groups to only `r (((filter(SVIVA2_01_comb,SYMBOL==2,MEMORY_score==2)%>%nrow())/(filter(SVIVA2_01_comb,SYMBOL==2,MEMORY_score>=0)%>%nrow()))*100)%>%round(0)`%.[^footnote_memory_symbols] In sum, with respect to both measures, the mean of the symbols condition was significantly lower compared with both control groups, and the differences between the two control groups are insignificant. In additional analyses, not shown here, I also account for the two policies, with no change to the results.

[^footnote_RT_symbols]: The differences in seconds as extracted from the differences in logs are: `r (x_realsymbol_RT_log-x_nosymbol_RT_log)%>%abs()%>%exp()%>%round(1)`-`r (x_realsymbol_RT_log-x_fakesymbol_RT_log)%>%abs()%>%exp()%>%round(1)`. Still, these differences are significant.

[^footnote_memory_symbols]: Because each participant was asked with regard to the last policy plan which was presented, the memory score is measured with regard to half of the observations (*n*=`r filter(SVIVA2_01_comb,MEMORY_score>=0)%>%nrow()`). The measure of memory test score is ordinal with three categories, and therefore I also tested it's correlation with the symbols manipulation via Spearman test. Spearman rho = `r cor_memory_nosymbol` and `r cor_memory_fakesymbol`, compared with the no symbols and fake symbols control groups.

**Figure 7: The effect of symbolic elements on elaboration**

```{r, echo=FALSE,warning=FALSE,message=FALSE, out.width=c('50%', '50%'),fig.show='hold',, fig.width=4,fig.height=4}
ggplot(SVIVA2_01_comb, aes(factor(SYMBOL_n), ELABORATION_time_log))+
  theme_classic()+
  #geom_boxplot(width=0.25,color = "gray70")+
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0.1,size=0.5,color="#9999CC",alpha=0.7) + 
  stat_summary(fun.y = mean, geom = "point", size=1.5)+
  stat_summary(fun.y = mean, aes(group = 1), geom = "line", size=0.5)+
  scale_x_discrete(name="",labels=c("no symbols","fake symbols","real symbols"))+
  scale_y_continuous(name = "Reaction Time (log)")

ggplot(SVIVA2_01_comb, aes(factor(SYMBOL_n), MEMORY_score))+
  theme_classic()+
  #geom_boxplot(width=0.25,color = "gray70")+
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width=0.1,size=0.5,color="#9999CC",alpha=0.7) + 
  stat_summary(fun.y = mean, geom = "point", size=1.5)+
  stat_summary(fun.y = mean, aes(group = 1), geom = "line", size=0.5)+
  scale_x_discrete(name="",labels=c("no symbols","fake symbols","real symbols"))+
  scale_y_continuous(name = "Memory score")
```

The less subjects elaborate on the weak policy plans, the more they are likely to report higher levels of trust in them, which will be closer to the levels of trust in the strong policy. To test mediation mechanism, I used the causal mediation analysis method via `mediate` R package, which was developed by Imai et al. [@imai_2010; @imai_2011]. This method predicts the change in the outcome variable (here: trust in policy) given the causal change in the mediator that would be induced by the treatment (here: decreased elaboration inducted by the real symbols). Accordingly, it enables to decipher the proportion of the causal mediation effect from the total effect. This method is based on counter-factual comparisons, as opposed to the traditional method of @baron_1986, which is based on linear structural equation modeling. It is advantaged in terms of its accuracy as well as its ability to identify causality. The results of the causal analysis tables are summarized in **Table 3**.

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(mediation)
set.seed(2018)

#RT
med.fit_RT_nosymbol = 
  lm(ELABORATION_time_log~SYMBOL*INFORMATION,
     data=filter(SVIVA2_01_comb,SYMBOL!=1))

med.out_RT_nosymbol = 
  lm(trust~SYMBOL*INFORMATION+ELABORATION_time_log*INFORMATION,
     data=filter(SVIVA2_01_comb,SYMBOL!=1))

med.fit_RT_fakesymbol = 
  lm(ELABORATION_time_log~SYMBOL*INFORMATION,
     data=filter(SVIVA2_01_comb,SYMBOL!=0))

med.out_RT_fakesymbol = 
  lm(trust~SYMBOL*INFORMATION+ELABORATION_time_log*INFORMATION,
     data=filter(SVIVA2_01_comb,SYMBOL!=0))

med_RT__nosymbol_INFORMTION_0 = mediate(med.fit_RT_nosymbol,
                                med.out_RT_nosymbol,
                                treat = "SYMBOL",
                                mediator = "ELABORATION_time_log", 
                                covariates = list(INFORMATION = 0), sims = 100)

med_RT__fakesymbol_INFORMTION_0 = mediate(med.fit_RT_fakesymbol,
                                  med.out_RT_fakesymbol,
                                  treat = "SYMBOL",
                                  mediator = "ELABORATION_time_log", 
                                  covariates = list(INFORMATION = 0), sims = 100)

med_RT__nosymbol_INFORMTION_1 = mediate(med.fit_RT_nosymbol,
                                med.out_RT_nosymbol,
                                treat = "SYMBOL",
                                mediator = "ELABORATION_time_log", 
                                covariates = list(INFORMATION = 1), sims = 100)

med_RT__fakesymbol_INFORMTION_1 = mediate(med.fit_RT_fakesymbol,
                                  med.out_RT_fakesymbol,
                                  treat = "SYMBOL",
                                  mediator = "ELABORATION_time_log", 
                                  covariates = list(INFORMATION = 1), sims = 100)

#memory
med.fit_memory_nosymbol = 
  lm(MEMORY_score~SYMBOL*INFORMATION,
     data=filter(SVIVA2_01_comb,SYMBOL!=1))

med.out_memory_nosymbol = 
  lm(trust~SYMBOL*INFORMATION+MEMORY_score*INFORMATION,
     data=filter(SVIVA2_01_comb,SYMBOL!=1))

med.fit_memory_fakesymbol = 
  lm(MEMORY_score~SYMBOL*INFORMATION,
     data=filter(SVIVA2_01_comb,SYMBOL!=0))

med.out_memory_fakesymbol = 
  lm(trust~SYMBOL*INFORMATION+MEMORY_score*INFORMATION,
     data=filter(SVIVA2_01_comb,SYMBOL!=0))

med_memory__nosymbol_INFORMTION_0 = mediate(med.fit_memory_nosymbol,
                                med.out_memory_nosymbol,
                                treat = "SYMBOL",
                                mediator = "MEMORY_score", 
                                covariates = list(INFORMATION = 0), sims = 100)

med_memory__fakesymbol_INFORMTION_0 = mediate(med.fit_memory_fakesymbol,
                                  med.out_memory_fakesymbol,
                                  treat = "SYMBOL",
                                  mediator = "MEMORY_score", 
                                  covariates = list(INFORMATION = 0), sims = 100)

med_memory__nosymbol_INFORMTION_1 = mediate(med.fit_memory_nosymbol,
                                med.out_memory_nosymbol,
                                treat = "SYMBOL",
                                mediator = "MEMORY_score", 
                                covariates = list(INFORMATION = 1), sims = 100)

med_memory__fakesymbol_INFORMTION_1 = mediate(med.fit_memory_fakesymbol,
                                  med.out_memory_fakesymbol,
                                  treat = "SYMBOL",
                                  mediator = "MEMORY_score", 
                                  covariates = list(INFORMATION = 1), sims = 100)

extract_mediation_summary <- function (x) {

  clp <- 100 * x$conf.level
  isLinear.y <- ((class(x$model.y)[1] %in% c("lm", "rq")) || 
                   (inherits(x$model.y, "glm") && x$model.y$family$family == 
                      "gaussian" && x$model.y$family$link == "identity") || 
                   (inherits(x$model.y, "survreg") && x$model.y$dist == 
                      "gaussian"))

  printone <- !x$INT && isLinear.y

  if (printone) {

    smat <- c(x$d1, x$d1.ci, x$d1.p)
    smat <- rbind(smat, c(x$z0, x$z0.ci, x$z0.p))
    smat <- rbind(smat, c(x$tau.coef, x$tau.ci, x$tau.p))
    smat <- rbind(smat, c(x$n0, x$n0.ci, x$n0.p))

    rownames(smat) <- c("ACME", "ADE", "Total Effect", "Prop. Mediated")

  } else {
    smat <- c(x$d0, x$d0.ci, x$d0.p)
    smat <- rbind(smat, c(x$d1, x$d1.ci, x$d1.p))
    smat <- rbind(smat, c(x$z0, x$z0.ci, x$z0.p))
    smat <- rbind(smat, c(x$z1, x$z1.ci, x$z1.p))
    smat <- rbind(smat, c(x$tau.coef, x$tau.ci, x$tau.p))
    smat <- rbind(smat, c(x$n0, x$n0.ci, x$n0.p))
    smat <- rbind(smat, c(x$n1, x$n1.ci, x$n1.p))
    smat <- rbind(smat, c(x$d.avg, x$d.avg.ci, x$d.avg.p))
    smat <- rbind(smat, c(x$z.avg, x$z.avg.ci, x$z.avg.p))
    smat <- rbind(smat, c(x$n.avg, x$n.avg.ci, x$n.avg.p))

    rownames(smat) <- c("ACME (control)", "ACME (treated)", 
                        "ADE (control)", "ADE (treated)", "Total Effect", 
                        "Prop. Mediated (control)", "Prop. Mediated (treated)", 
                        "ACME (average)", "ADE (average)", "Prop. Mediated (average)")

  }

  colnames(smat) <- c("Estimate", paste(clp, "% CI Lower", sep = ""), 
                      paste(clp, "% CI Upper", sep = ""), "p-value")
  smat

}

```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
#RT
table_med_RT__nosymbol_INFORMTION_0 = 
summary(med_RT__nosymbol_INFORMTION_0) %>% 
extract_mediation_summary()

table_med_RT__fakesymbol_INFORMTION_0 = 
summary(med_RT__fakesymbol_INFORMTION_0) %>% 
extract_mediation_summary()

table_med_RT__nosymbol_INFORMTION_1 = 
summary(med_RT__nosymbol_INFORMTION_1) %>% 
extract_mediation_summary()

table_med_RT__fakesymbol_INFORMTION_1 = 
summary(med_RT__fakesymbol_INFORMTION_1) %>% 
extract_mediation_summary()

#memory
table_med_memory__nosymbol_INFORMTION_0 = 
summary(med_memory__nosymbol_INFORMTION_0) %>% 
extract_mediation_summary()

table_med_memory__fakesymbol_INFORMTION_0 = 
summary(med_memory__fakesymbol_INFORMTION_0) %>% 
extract_mediation_summary()

table_med_memory__nosymbol_INFORMTION_1 = 
summary(med_memory__nosymbol_INFORMTION_1) %>% 
extract_mediation_summary()

table_med_memory__fakesymbol_INFORMTION_1 = 
summary(med_memory__fakesymbol_INFORMTION_1) %>% 
extract_mediation_summary()
```

**Table 3: Causal mediation analysis**

*Mediator: Reaction time*
3.1: Weak policy
```{r, echo=FALSE,warning=FALSE,message=FALSE}
table_med_RT_INFORMTION_0 = 
  cbind.data.frame(table_med_RT__nosymbol_INFORMTION_0,
                   table_med_RT__fakesymbol_INFORMTION_0)
library(kableExtra)

kable(table_med_RT_INFORMTION_0, "html",digits = 3) %>%
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = F, position = "left", font_size = 11) %>% 
  add_header_above(c(" " = 1, "real vs. no symbols" = 4, "real vs. fake symbols" = 4))
```

3.2: Strong policy
```{r, echo=FALSE,warning=FALSE,message=FALSE}
table_med_RT_INFORMTION_1 = 
  cbind.data.frame(table_med_RT__nosymbol_INFORMTION_1,
                   table_med_RT__fakesymbol_INFORMTION_1)

kable(table_med_RT_INFORMTION_1, "html",digits = 3) %>%
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = F, position = "left", font_size = 11) %>% 
  add_header_above(c(" " = 1, "real vs. no symbols" = 4, "real vs. fake symbols" = 4))

```

*Mediator: Memory score*
3.3: Weak policy
```{r, echo=FALSE,warning=FALSE,message=FALSE}
table_med_memory_INFORMTION_0 = 
  cbind.data.frame(table_med_memory__nosymbol_INFORMTION_0,
                   table_med_memory__fakesymbol_INFORMTION_0)

kable(table_med_memory_INFORMTION_0, "html",digits = 3) %>%
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = F, position = "left", font_size = 11) %>% 
  add_header_above(c(" " = 1, "real vs. no symbols" = 4, "real vs. fake symbols" = 4))
```

3.4: Strong policy
```{r, echo=FALSE,warning=FALSE,message=FALSE}
table_med_memory_INFORMTION_1 = 
  cbind.data.frame(table_med_memory__nosymbol_INFORMTION_1,
                   table_med_memory__fakesymbol_INFORMTION_1)

kable(table_med_memory_INFORMTION_1, "html",digits = 3) %>%
  kable_styling(bootstrap_options = c("striped","condensed"), full_width = F, position = "left", font_size = 11) %>% 
  add_header_above(c(" " = 1, "real vs. no symbols" = 4, "real vs. fake symbols" = 4))
```

*Notes*: "ACME": Averaged causal mediated effect; "ADE": Averaged direct effect. Mediate function uses bootstrap to estimate the confidence intervals. In all these analyses I used `set.seed=2018`.

The results suggest that the decreased reaction time significantly mediates the effect of the real symbols on trust in the weak policies. The estimated proportion of the mediation is `r (table_med_RT__nosymbol_INFORMTION_0["Prop. Mediated","Estimate"]*100) %>% round(1)`% compared with the no symbols, and `r (table_med_RT__fakesymbol_INFORMTION_0["Prop. Mediated","Estimate"]*100)%>%round(1)`% compared with the fake symbols. The decreased memory score also mediates the effect (*p*<.1), but only compared with the no symbols group.[^footnote_mediation_memory] The effect of the real symbols on trust in the strong policies is not significantly mediated by the measures of elaboration. Altogether, these analyses are consistent with my H4, and suggest that subjects who saw the policy plans with the familiar symbols were less likely to scrutinize their content, which in turn increased their trust in the weak policies.

[^footnote_mediation_memory]: The differences between the analyses of the two measures can be also explained by the fact that memory score is calculated with regard to half of the observations.

<br>

#Discussion and conclusion

A growing body of public administration research seeks to assess whether and how government organizations can increse citizens' trust through direct communications with citizens. This study contributes to this discussion, by pointing to the influence of symbolic elements in public communications, and unpacking the theoretical mechanisms underlying their effect on citizens' trust. Using a rigorous experimental design, embedded in a realistic setting of communication of policy plans of the Israeli Environmental Protection Ministry, I tested the effect of symbolic elements on trust, and its interactions with the substantive content of the communications and with their perceived personal relevance. The experimental results demonstrate that using familiar symbolic elements in communications about policy plans increase citizens' trust in these policies, via two mechanisms: First, they evoke positive fellings that are transferred to the policy plan. Second, by putting citizens in a positive mood, they distract them from thinking about the substantive content of the communication, and thus - they cause them to disregard its weaknesses, if exist. the latter mechanism entails an interaction between symbols and weak information. These theoretical mechanisms can be understood in light of the social psychology elaboration likelihood model. In addition, I theorized that the effect of symbols is attenuated the more the communication is personally relevant for citizens' lives, but the reults do not support this expectation. Rather, I found that citizens who perceived the policy issue as highly relevant for their lives (i.e. air-pollution in their area of residence), were similarly distracted by the graphic symbols, and persuaded to trust governmnet's actions, even when they were undoubtedly illogical. These findings are based on a large sample of more than `r (SVIVA2_01%>%nrow() %>% round(-2))-100` participants, and are robust for various model specifications.

What are the implications of these findings for public administation resarch? First, they highlight the important role of strategically designed symbolic elements in communications in shaping citizens' attitudes about government. Citizens' responses to symbolic elements and branding thus far received relatively little attention in Public administration literature, compared with the the abounded research on responses to information. Yet, this study, and others, demonstrated that symbols can be fairly effective in increasing citizens' trust, and mitigating their negative attitudes [@alonbarkat_2017; @karens_2016; @teodoro_2018]. Furthermore, this study demostrates that symbols decrease the "rational", systematic processing of substantive information. This may suggest that government organizations may bypass the cognitive biases that often undermine the effectiveness of positive information, by employing symbolic communication.[^footnote_biases_ELM] This is good news as far as we are concerned mainly about citizens' *undervaluation* of government performance. On the other hand, it challenges the democratic ideal that citizens' form their attitudes about governments on the basis of rational thinking [###Dahl...]. Moreover, it suggests that public organizations may use symbols in communications to compensate for their poor performance and inefficient policy plans. In turn, by preserving citizens' trust in them and escaping from public blame, organizations can nurture their long-term positive reputation, which attenuates politicians' incentives to monitor and oversight them [##Busuioc_2016; Bertelli_2018##]. Overall... something about democratic accountability...

[^footnote_biases_ELM]: This posibility is supported, theoretically, by ELM. According to ELM, biased processing occures in the central route, whereby people give unproportional weight for strong or weak arguments. Accordingly, decresaing the ammount of thinking may also decrease biased processing [@petty_1986]. Future research might put this expectation to test.

Methodologically, these findings highlight that it is important to test the effectiveness of public communications, and not only the information.

Almost any government communication involves design elements, and many of them include familiar symbols and brand elements. Conversely, some of the current studies that explored the effect of information, did so by comparing participants' responses to pieces of information (mainly favorable vs. unfavorable performance) which were presented in surveys in the form of plain text and data tables without design elements which would be normally included in  such communications [@baekgaard_2016; @hvidman_2016; ]. This questions our ability to generalize from these experiments to citizens' responses to public communications in the real world.

Finally, PA can also implement ELM...

How can we explain the null findings about personal relevance %>%

limitations...

<br>

##Appendix A: Information manipulation

**Air-pollution policy** (Reducing the air-pollution in the Haifa-Bay)

EPM is working to reduce the air-pollution in the Haifa-Bay through:

*Strong*

* **Increasing the supervision of the factories in Haifa-Bay**.
In order to reduce the emissions of pollutants from the factories in the industrial area, the Ministry is working to increase the supervision of the factories in several ways. The ministry will increase the number of inspection visits, expand the use of unannounced inspections, and increase the amount of fines for pollutants.

* **Reduction of smoke emissions from vehicles in Haifa Bay**.
The Ministry is taking a number of actions to reduce the amount of air-pollution from vehicles in the area. The ministry promotes the installation of particle filters in hundreds of vehicles in the area, and promotes the operation of a "clean-air area" that restricts the entry of polluting vehicles.

*Weak*

* **Decreasing the supervision of the factories in Haifa-Bay**.
The Ministry is preparing for a change in the method of supervision of the factories. The ministry will reduce the number of inspection visits and unannounced inspections, rely on factories' self-reports on pollutant emissions, and reduce the amount of fines. The change is coordinated with members of the industry, which are committed to reducing the pollution, and it will assist to build mutual respect between them and the ministry.

* **Raising citizens' awareness of the efforts to improve air quality**.
In recent years, the Ministry has taken steps to reduce air pollution in Haifa-Bay, which have been partially successful. The Ministry is working to raise awareness of its efforts and to emphasize their success in order to improve its image among the residents. To this end, the ministry has increased its investment in public relations and publicity in this subject.

<br>

**Recycling policy** (Reducing waste and increasing recycling)

EPM is working to reduce the amount of waste and increase its recycling through:

*Strong*

* **Packaging recycling - Increased supervision of manufacturers**.
In order to reduce the amount of waste generated from packaging and to encourage recycling, the Ministry implements the Packaging act, which imposes on manufacturers and importers the responsibility for recycling the packaging. The Ministry will work to increase enforcement by virtue of this act, and will enforce it with regard to additional manufacturers and importers.

* **Reducing the usage of disposable bags**.
In order to reduce the use of disposable plastic bags, the Ministry enforces the Bags act, which sets a compulsory charge on bags sold in large retail chains and imposes reporting obligations on the chains. The Ministry works to implement the act, and encourages the transition to reusable grocery bags.

*Weak*

* **Packaging recycling - Decreased supervision of manufacturers**.
The Ministry is preparing to a change in the method of supervision of the manufacturers, who are responsible for the recycling of product packaging by virtue of the act The ministry will reduce the number of inspections, rely on manufacturers' self-reports regarding their compliance with the requirements of the act, and reduce the amount of fines. The change is coordinated with members of the industry, which are committed to the issue, and it will assist to strengthen mutual respect between them and the ministry.

* **Reducing the usage of disposable bags**.
In recent years, the Ministry has taken steps to encourage the reduction of waste, which have been partially successful. The Ministry works to raise awareness of the actions taken and to highlight their success in order to improve its image among the public. To this end, the ministry has increased its investment in public relations and publicity in this subject.

<br>

##Appendix B: Symbols manipulation
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/symbols_manipulation.png)

<font size="0.5">

*Note: The policy plans communications presented here are the strong policies, in the version adjusted for mobile interface.*

</font>

<br>

<br>

#References